---
title: "Practice Activity 3: Data Wrangling with dplyr 🏫"
format: html
editor: visual
---

Today you will be using the dplyr package to clean a dataset and then using that cleaned dataset to figure out what college Ephelia has been accepted to.

# Part One: Data import and cleaning

This section will clean today's dataset, so that you can use it more easily in Part Two.

## Data download & packages

First, we declare our package dependencies and load the data.

(Note that the data loading function `read_csv()` will give you an outpouring of helpful information about the dataset. If you do not see the word "error", there is nothing to be concerned about.)

```{r setup}
#| warning: false
#| message: false

library(tidyverse)

colleges <- read_csv(here::here("supporting_artifacts", 
                              "Practice_Activities", 
                              "Week3",
                              "colleges.csv")
                    )
```

## Data cleaning

Now we will clean the data. Alas, each of the R chunks in this section will cause an error and / or do the desired task incorrectly. Even the chunks that run without error **are not** correct! You will need to find the mistake and correct it to complete the intended action.

**Step 1:** There are too many variables in this dataset. We don't need all of them. Narrow your dataset down to only:

-   Name of the institution
-   City, State, and ZIP code of the institution
-   The Admissions Rate
-   The average SAT score
-   The number of undergraduate students
-   The in and out of state tuition
-   Whether the school is public or private
-   The "REGION" variable.

```{r select-vars}
colleges_clean <- colleges |> 
  select(INSTNM, CITY, STABBR, ZIP,
         ADM_RATE, SAT_AVG, UGDS,
         TUITIONFEE_IN, TUITIONFEE_OUT,
         CONTROL, REGION) 
```

**Step 2:** Remove the schools that are private and for-profit (category 3).

```{r filter-private-profit}
colleges_clean <- colleges_clean |> 
  filter(CONTROL < 3) 
```

**Step 4:** Adjust the appropriate variables to be numeric.

```{r mutate-numeric}
#| message: false
colleges_clean <- colleges_clean |> 
  mutate(
    TUITIONFEE_IN = as.numeric(TUITIONFEE_IN),
    TUITIONFEE_OUT = as.numeric(TUITIONFEE_OUT),
    SAT_AVG = as.numeric(SAT_AVG),
    ADM_RATE = as.numeric(ADM_RATE)
    ) 
```

NOTE: The NAs introduced are due to blanks in the dataset.

**Step 5:** Adjust the appropriate variables to be factors.

```{r mutate-factor}
colleges_clean <- colleges_clean |>
  mutate(
    CONTROL = as.factor(CONTROL),
    REGION = as.factor(REGION)
)
```

**Step 6:** Create a new variable called TUITION_DIFF which contains the difference between in and out of state costs.

```{r mutate-differences}
colleges_clean <- colleges_clean |> 
    mutate(TUITION_DIFF = TUITIONFEE_OUT - TUITIONFEE_IN)
```

**Step 7:** Remove every row with missing data.

***Note*****:** This is not always a great idea! Usually, even if *some* of the information is missing, we don't want to throw out the entire row. This time, however, we'll be lazy.

```{r drop-na}
colleges_clean <- colleges_clean |> 
  drop_na()
```

Lastly, notice that each of these steps started with

```{r all-together}
#| eval: false

colleges_clean <- colleges_clean |> ...
```

That is pretty redundant! Instead, we could perform all these tasks as one long "pipeline."

**Step 8:** Combine your (fixed) code chunks into a **single** code chunk that carries out all of the steps necessary to clean the data.

**Note:** Think about coding efficiency -- you **should not** have multiple calls to the **same** function!

```{r combine}
#| message: false
#| warning: false
# Code combining ALL of your previous steps into ONE pipeline

colleges_clean <- colleges |> 
  select(INSTNM, CITY, STABBR, ZIP,
         ADM_RATE, SAT_AVG, UGDS,
         TUITIONFEE_IN, TUITIONFEE_OUT,
         CONTROL, REGION) |> 
  filter(CONTROL < 3) |> 
  mutate(
    TUITIONFEE_IN = as.numeric(TUITIONFEE_IN),
    TUITIONFEE_OUT = as.numeric(TUITIONFEE_OUT),
    SAT_AVG = as.numeric(SAT_AVG),
    ADM_RATE = as.numeric(ADM_RATE)
    ) |> 
  mutate(
    CONTROL = as.factor(CONTROL),
    REGION = as.factor(REGION)
    ) |> 
  mutate(TUITION_DIFF = TUITIONFEE_OUT - TUITIONFEE_IN
         ) |> 
  drop_na() 


colleges_clean <- colleges |> 
  select(INSTNM, CITY, STABBR, ZIP,
         ADM_RATE, SAT_AVG, UGDS,
         TUITIONFEE_IN, TUITIONFEE_OUT,
         CONTROL, REGION) |> 
  filter(CONTROL < 3) |> 
  mutate(across(c(TUITIONFEE_IN, TUITIONFEE_OUT,
                  SAT_AVG, ADM_RATE),
                as.numeric),
         across(c(CONTROL, REGION),
                as.factor),
         TUITION_DIFF = TUITIONFEE_OUT - TUITIONFEE_IN
         ) |> 
  drop_na() 
```

# Part Two: Identify the mystery college

Wow! Your best friend Ephelia has been accepted to her top choice college! Unfortunately, Ephelia is a very mysterious person, and she won't tell you directly which college this is. You'll have to use her clues to figure out which school is her dream school.

Clues:

1.  This college is located in Region 7.

2.  This college's admission rate is above the median rate for the region.

3.  This college **does not** charge the same for in- and out-of-state tuition.

4.  The average SAT score of this college is an odd number.

5.  This college **is not** in Idaho.

6.  Less than 1,000 people apply to this college every year. (Assume the size of the first year class is 1/4 of the undergraduate population.)

7.  The college **is not** in the state where Dr. Theobold received a Ph.D.

8.  Of the three options remaining at this step, Ephelia will attend the cheapest one.

undergrads

```{r}
colleges_clean |> 
  filter(REGION == 7) |> 
  filter(ADM_RATE > median(ADM_RATE)
         ) |> 
  filter(TUITION_DIFF > 0) |> 
  filter(SAT_AVG %% 2 == 1) |> 
  filter(STABBR != "ID") |> 
  mutate(UGDS = as.numeric(UGDS)
         ) |> 
  mutate(FRESH = UGDS / 4) |> 
  mutate(APP_FRESH = FRESH / ADM_RATE) |> 
  filter(APP_FRESH < 1000) |> 
  filter(STABBR != "MT") |> 
  slice_min(TUITION_DIFF)
  

colleges_clean |> 
  filter(REGION == 7,
         ADM_RATE > median(ADM_RATE),
         TUITION_DIFF > 0,
         SAT_AVG %% 2 == 1, 
         STABBR != "ID") |> 
  mutate(UGDS = as.numeric(UGDS),
         FRESH = UGDS / 4,
         APP_FRESH = FRESH / ADM_RATE) |> 
  filter(APP_FRESH < 1000,
         STABBR != "MT") |> 
  slice_min(TUITION_DIFF)
  
  
```

**freshman = ugds / 4**

**applicaiton = freshman / admrate** 2000/0.5

make 2 new variables in a mutate step
